---
title: "R Notebook"
output: 
  html_notebook:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  # if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---
# Loading Data and Univariate Analysis
```{r}
otrain <- read_csv('data/train.csv')
otest <- read_csv('data/test.csv')
glimpse(otrain)
```

```{r}
summary(otrain$application_date)
```

```{r}
table(otrain$segment, useNA = "ifany")
```

```{r}
sort(table(otrain$branch_id, useNA = 'ifany'))
n_distinct(otrain$branch_id)
```

```{r}
sort(round(prop.table(table(otrain$state, useNA = 'ifany'))*100, 2))
n_distinct(otrain$state)
```

```{r}
sort(round(prop.table(table(otrain$zone, useNA = 'ifany'))*100, 2))
```

```{r}
summary(otrain$case_count)
```

# MultiVariate Analysis
## Zone -> State -> Branch
Segment 2 has NAs in branch and zone. But, we can get zone from state using zone-state mapping for segment 1. But we can't get branch that way
```{r}
zone_state_mapping <- otrain %>% 
  dplyr::filter(segment == 1) %>% 
  select(zone, state) %>% 
  distinct() %>% 
  arrange(zone)

train_preped <- otrain %>% 
  left_join(zone_state_mapping, by = "state", suffix = c("_old", "")) %>% 
  select(-zone_old)

sort(round(prop.table(table(train_preped$zone, useNA = 'ifany'))*100, 2))
```

```{r}
# state_zone_table <- addmargins(round(prop.table(table(train_preped$state, train_preped$zone))*100, 2))
# state_zone_table[order(state_zone_table[,"Sum"], decreasing = TRUE),]
```

```{r}
train_preped <- train_preped %>% 
  mutate(segment = paste('s', segment, sep = '_'),
         branch_id = paste('b', branch_id, sep = '_'))

train_preped$pathString <- paste(
  "country",
  train_preped$segment,
  train_preped$zone,
  train_preped$state,
  train_preped$branch_id,
  sep = "/"
)

train_hierarchy <- as.Node(train_preped)
```

```{r}
# print(train_hierarchy, limit = 20)

# Prune(population, pruneFun = function(x) !x$isLeaf || x$case_count > 100)

train_network <- ToDataFrameNetwork(train_hierarchy, direction = "climb")
simpleNetwork(train_network, fontSize = 12)
```

## With target
```{r}
train_wide <- train_preped %>% 
  pivot_wider(id_cols = application_date, names_from = c(zone, segment), values_from = case_count, values_fill = list(case_count = 0), values_fn = list(case_count = sum))

sapply(train_wide, function(x) sum(x==0, na.rm = TRUE))

lapply(train_wide, summary)
```

```{r}
train_wide_xts <- xts(dplyr::select(train_wide, -application_date), order.by = train_wide$application_date)

head(train_wide_xts)
```

### Segment
```{r}
dygraph(train_wide_xts[,c("EAST_s_1", "EAST_s_2")]) %>% 
  dySeries("EAST_s_1", axis = 'y', color = "blue") %>% 
  dySeries("EAST_s_2", axis = 'y2', color = "green")
```

### Zone
```{r, fig.width=10, fig.height=4}
train_wide_xts[,c("EAST_s_1", "NORTH_s_1", "SOUTH_s_1", "WEST_s_1", "CENTRAL_s_1")] %>% 
  dygraph(main = "Count of Applications over Time") %>% 
  dyOptions(drawPoints = TRUE, pointSize = 1, colors = RColorBrewer::brewer.pal(5, "Set2")) %>% 
  dySeries("EAST_s_1", axis = 'y2') %>%
  dySeries("NORTH_s_1", axis = 'y') %>%
  dySeries("SOUTH_s_1", axis = 'y2') %>%
  dySeries("WEST_s_1", axis = 'y2') %>%
  dySeries("CENTRAL_s_1", axis = 'y') %>%
  dyRangeSelector() %>% 
  dyHighlight(highlightCircleSize = 2, 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = TRUE)
```

### Segment + Zone
```{r}
ggplot(train_preped, aes(x = application_date, y = case_count)) +
  geom_line() +
  facet_wrap( ~ segment + zone, scales = "free_y", ncol = 5)
```

### State
```{r}
# ggplotly(
#   train_preped %>% 
#   dplyr::filter(segment == "s_1", zone == "NORTH") %>% 
#   ggplot(aes(x = application_date, y = case_count)) +
#   geom_line(aes(color = as.factor(state))) +
#   theme_minimal()
# )

ggplotly(
  train_preped %>%
    dplyr::filter(segment == "s_1", zone == "NORTH") %>%
    ggplot(aes(x = application_date, y = case_count)) +
    geom_area(aes(color = state, fill = state),
              alpha = 0.5,
              position = position_dodge(0.8)) +
    stat_smooth(method = "loess", aes(color = state, fill = state)) +
    theme_minimal()
)
```

### Branch
```{r}
ggplotly(
  train_preped %>%
    dplyr::filter(segment == "s_1", zone == "NORTH", state == "PUNJAB") %>%
    ggplot(aes(x = application_date, y = case_count)) +
    geom_area(
      aes(color = branch_id, fill = branch_id),
      alpha = 0.5,
      position = position_dodge(0.8)
    ) +
    stat_smooth(method = "loess", aes(color = branch_id, fill = branch_id)) +
    scale_x_date(date_labels = "%b/%Y", date_breaks = "1 month") +
    theme(axis.text.x = element_text(angle=45, hjust = 1))
)
```

### Day of month
Adding date related info.
```{r}
train_preped <- train_preped %>% 
  mutate(day_of_month = mday(application_date),
         day_of_week = wday(application_date, label = TRUE, abbr = TRUE),
         is_end_of_month = (application_date == (ceiling_date(application_date, unit = "months") - days(1))),
         month = month(application_date, label = TRUE, abbr = TRUE),
         year = year(application_date),
         year_month = paste(year, month, sep = "_"))
```

Dividing by segment
```{r}

train_preped_seg1 <- train_preped %>% 
  dplyr::filter(segment == "s_1")

train_preped_seg2 <- train_preped %>% 
  dplyr::filter(segment == "s_2")
```

```{r}
ggplotly(
  train_preped_seg1 %>%
    dplyr::filter(day_of_month %in% c(1,31)) %>% 
    ggplot(aes(x = application_date, y = case_count)) +
    geom_line(aes(color = as.factor(day_of_month)),
              alpha = 0.5,
              position = position_dodge(0.8)) +
    theme_minimal()
)
```

```{r}
train_preped_seg1 %>% 
  group_by(day_of_month) %>% 
  group_map(~summary(.x$case_count))

train_preped_seg1 %>% 
  group_by(day_of_month) %>% 
  group_map(~sd(.x$case_count))
```

```{r}
ggplotly(
  train_preped_seg1 %>%
    ggplot(aes(x =  case_count)) +
    geom_density(aes(color = as.factor(day_of_month), fill = as.factor(day_of_month)),
              alpha = 0.5,
              position = position_dodge(0.8)) +
    theme_minimal()
)
```

```{r}
ggplotly(
  train_preped_seg1 %>%
    group_by(day_of_month) %>% 
    mutate(case_count_quantile_max = quantile(case_count, probs = .9)) %>% 
    dplyr::filter(case_count <= case_count_quantile_max) %>% 
    ungroup() %>% 
    ggplot(aes(x =  case_count)) +
    geom_density(aes(color = as.factor(day_of_month), fill = as.factor(day_of_month)),
              alpha = 0.5,
              position = position_dodge(0.8)) +
    theme_minimal()
)
```
### Day of week
```{r}
train_preped_seg1 %>% 
  group_by(day_of_week) %>% 
  group_map(~summary(.x$case_count))

train_preped_seg1 %>% 
  group_by(day_of_week) %>% 
  group_map(~sd(.x$case_count))
```

```{r}
ggplotly(
  train_preped_seg1 %>%
    ggplot(aes(x = case_count)) +
    geom_density(aes(color = day_of_week, fill = day_of_week),
              alpha = 0.3,
              position = position_dodge(0.8)) +
    theme_minimal()
)
```

### why day of month varies
Check by month, is_end_of_month, year_month, zone, state, 
```{r}
ggplotly(
  train_preped_seg1 %>%
    dplyr::filter(day_of_month == 31) %>% 
    ggplot(aes(x =  case_count)) +
    geom_density(aes(color = year_month, fill = year_month),
              alpha = 0.3,
              position = position_dodge(0.8)) +
    theme_minimal()
)
```













